---
title: "Pythium populations in Michigan greenhouses"
output:
  html_document:
    keep_md: yes
editor_options:
  chunk_output_type: console
---

```{r custom_functions, echo=FALSE, message=FALSE}
# ipak function: install and load multiple R packages.
# check to see if packages are installed. Install them if they are not, then load them into the R session.
# Source: https://gist.github.com/stevenworthington/3178163
ipak <- function(pkg){
new.pkg <- pkg[!(pkg %in% installed.packages()[,"Package"])]
if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE, repos = "https://cran.rstudio.com")
sapply(pkg, require, character.only = TRUE)
}
```


```{r libraries, echo=FALSE, message=FALSE, results='hide'}
packages <- c("ggplot2","RColorBrewer","poppr", "dplyr", "knitr","kableExtra", "here")
ipak(packages)
```

## Data input

```{r}
#Reading genealex object
ultimhier <- read.genalex(here("data", "ult_pop_clean.csv"))

#recode populations

# ultimhier@pop <- dplyr::recode(ultimhier@pop, 
#                                "d_ff"="wayne_fall-12", 
#                                "d_ss"="wayne_spring-13", 
#                                "g_f"="kent_fall-11", 
#                                "k_f"="kalamazoo_fall-11", 
#                                "k_ff"="kalamazoo_fall-12",
#                                "k_ss"="kalamazoo_spring-13")

splitStrata(ultimhier, sep = "_") <- ~County/Season
```


## Genotype accumulation curve

The aim is to determine the if the SSRs are enough to describe the genotypes in the population.

```{r locus_genotype}
#Locus table
locus_table(ultimhier)

#genotype accumulation curve
gac <- genotype_curve(ultimhier, sample = 1000, quiet = TRUE, thresh = 0.9)
```

## Genotype diversity

Basic summary of Pythium population data and definition of multilocus genotypes.

```{r summary}
setPop(ultimhier)<- ~Season
#Table
ult_table <- poppr(ultimhier)
ult_table$Simpson <- ult_table$lambda*(ult_table$N/(ult_table$N - 1))

knitr::kable(ult_table[,-13], format="markdown", digits = 3,
             caption=" Summary statistics of _P. ultimum_ populations")
```

Generating a table for the multilocus genotypes across populations
```{r mlgs, fig.width=10}
#Matrix indicating mulitlocus genotypes (MLGs) by population
ult.tab <- mlg.table(ultimhier)
#Associating MLGs to individual names
mlg.ult <- mlg.id(ultimhier)
#write.table(ult.tab, sep = ",", file = "data/mlg_tab.csv")
```


## Hierarchy and strata by factors

First, the strata will be recoded to facilitate interpretation of the data.  Then, the strata will be divided to generated a nested 
hierarchy.  The idea is that season is nested within county.

```{r}
#Accessing strata data of the genind object
ult.str <- strata(ultimhier)
knitr::kable(head(ult.str, n = 20), format = "markdown") 
```

Using strata, the populations could be defined using the strata.  A hierarchy using `county/season` or just by `season` or `county`.

```{r County_season}
setPop(ultimhier) <- ~County/Season
ultimhier

setPop(ultimhier)<- ~Season
ultimhier
```

If the populations are defined by **county**, there are three populations.  By looking at multilocus genotypes at the county level, Kalamazoo contains most of the multilocus genotypes.

```{r county_pop, fig.width=10}
#Setting up a population by county (3 populations/counties)
setPop(ultimhier) <- ~County

#Summary of multilocus genotypes by county
ult.tab <- mlg.table(ultimhier)

#write.table(poppr(ultimhier), sep = ",", file = "~/Documents/genotypebycounty.csv")
#write.table(poppr(ult.tab, sep = ",", file = "~/Documents/MLGbycounty.csv")
```

Now, if the poplation is defined by **season**, there will be three populations as well, these refer to seasons across years sampled.

```{r 3pop_data, fig.width=10}
setPop(ultimhier)<- ~Season
ult.tab <- mlg.table(ultimhier)
#write.table(ult.tab, sep = ",", file = "~/Documents/frequencies3pop.csv")
```

```{r, message=FALSE, warning=FALSE, fig.height=10}
setPop(ultimhier) <- ~County/Season
ult.tab.strata <- mlg.table(ultimhier, plot = FALSE)

ult.mlg.table <- as.data.frame(ult.tab.strata) %>% 
  dplyr::add_rownames(var = "County") %>% 
  tidyr::gather("MLG","n",2:66) %>%
  tidyr::separate("County",c("County", "Season"), sep ="_") %>%
  #group_by(County, Season, MLG) %>% summarise(n = sum(n)) %>%
  dplyr::filter(n > 0)


ggplot(ult.mlg.table, aes(x = MLG, y = n)) + 
  geom_bar(aes(fill = County), stat = "identity", color = "black") +
  scale_fill_manual(values = c("#252525","#ffffff", "#969696")) +
  theme_linedraw() + xlab("MLG") +
  theme(panel.grid.major.x = element_blank(), 
          panel.grid.minor.x = element_blank(),
          axis.text.x = element_text(size = 10, angle = 90, 
                                     hjust = 1, vjust = 0.5)) +
   facet_wrap(~Season, scales = "free_x", shrink = TRUE, 
              drop = TRUE, ncol = 1,
              labeller = labeller(Season = c(`fall-11`="Fall 11", `fall-12`="Fall 12", `spring-13`="Spring 12-13")))
  
#write.table(ult.mlg.table, sep = ",", file = "data/mlg_tab.csv")
```

### Shared MLG by Season

```{r}
mlg.crosspop(ultimhier, strata = ~Season)
```

### Shared MLG by County

```{r}
mlg.crosspop(ultimhier, strata = ~County)
```


```{r saving_rda, echo=FALSE, message=FALSE}
#save(ultimhier, file = "Pyult.rda")
```